Point transect distance sampling in inlabru

### 2019-08-10 paper.tex
Update paper.tex
Many edits, see email.

% RJC 09/08/2019 Added co-authors and affiliations. Note that I used \affil[]{} instead of \address[]{}
\author[1,*]{Andrew E Seaton}
\author[1,2]{Richard J Camp}
\author[1]{Janine B Illian}
\author[1]{David L Borchers}
\author[1]{David L Miller}      % Ask about co-authoring the manuscript
\author[1]{Len Thomas}          % Ask about co-authoring the manuscript
\author[1]{Stephen T Buckland}  % Ask about co-authoring the manuscript
\author[3]{Steve J Kendall}     % It is Steve's data we are using.

%\address{Centre for Research into Ecological \& Environmental Modelling and School of Mathematics \& Statistics, University of St Andrews, St Andrews, Fife, Scotland}
\affil[1]{Centre for Research into Ecological \& Environmental Modelling and School of Mathematics \& Statistics, University of St Andrews, St Andrews, Fife, Scotland}
\affil[2]{U. S. Geological Survey, Pacific Island Ecosystems Research Center, P.O. Box 44, \hawaii{} National Park, HI 96718, U.S.A.}
\affil[3]{U. S. Fish and Wildlife, Big Island National Wildlife Refuge Complex, 60 Nowelo St., Suite 100, Hilo, HI  96720, U.S.A.}
\affil[*]{Correspondence: Andrew E Seaton, Email: aes22@st-andrews.ac.uk}

@@

Introduction outline:
\begin{enumerate}
	\item Estimation of abundance and spatial distribution of animals is an important topic
	\item Distance sampling is a family of methods that aim to do this whilst accounting for imperfect detectability of animals
	\item Distance sampling began life under a design-based inference paradigm
	\item Randomised survey design is used in two ways.  Firstly, to justify the assumption that the true distribution of animals is uniform with respect to distance from the observer.  Secondly, to construct Horvit-Thompson-like estimators of animal density, which are then converted to abundance estimates for regions of interest
	\item Interest in the spatial distribution of animals has led to the development of approaches that combine traditional distance sampling methods and spatially explicit generalized additive models
	\item In this approach the estimated detection probabilities for discrete spatial units are used as an offset in a generalized additive model where the response variable is a count of detections within each unit
	\item This two-stage approach needs a way to propogate uncertainty from the detection model to the spatial model
	\item Recently [[JOYCE PAPER]] presented a one-stage approach that simultaneously estimates the detection and spatial processes, avoids binning data into counts and removes the need to consider uncertainty propagation
	\item This approach considered line-transect distance sampling as thinned log-Gaussian Cox process
	\item To estimate the detection function [[JOYCE et al]] constructed a spline-like function using an SPDE approach and fit the model using \texttt{R-INLA}
	\item In this paper we show how to achieve a similar one-stage model for point transect distance sampling
	\item Point transect surveys require a different approach than line transects because the area surveyed increases with distance from observer
	\item In addition to accounting for this we present an alternative way to specify the detection function.  We replace the spline approach with a parametric family of detection functions such as half-normal and hazard rate functions
	\item  These detection functions are typically not linear in their parameters and so cannot be straightforwardly estimated using methods for additive linear predictors
	\item To solve this we implement an iterative approach based on a Taylor approximation of the detection function and fit the model using the \texttt{inlabru} package which extends the functionality of \texttt{R-INLA}
	\item This represents a novel development in distance sampling methods that allows simultaneous estimation of parametric detection functions and the spatial distribution of animals from point transect data
\end{enumerate}

The estimation of the size of wild populations of animals is a critical objective within ecology and conservation [[citation]].  Distance sampling methods aim to achieve this by using a spatially explicit sampling design and estimating the detectability of animals as a function of distance from observer \citep{buckland_advanced_2008, buckland_distance_2015}.  Standard distance sampling approaches use a hybrid of design and model-based inference to estimate population size.  Probability of detection is modelled as a function of distance from observer and a randomized sampling design allows the construction of Horvitz-Thompson-like estimators of animal density \citep{horvitz_generalization_1952,  buckland_advanced_2008}.

Over time interest grew in fully model-based approaches which allow the use of non-randomized survey designs by considering a spatial model for animal density so observations can be associated with spatially indexed covariates and abundance can, in principle, be estimated for any subregion within the study area \citep{johnson_model-based_2010, buckland_model-based_2016, miller_spatial_2013}.  Model-based distance sampling has typically been implemented in a two-stage modelling process.  In the first stage, detectability is estimated within each sampling unit.  In the second stage, detections within sampling units are binned into counts and used as the response variable in generalized additive model framework where the detectability estimates from the first stage are used as an offset.  Due to often sparse nature of wildlife survey data this may require consideration of over or under-dispersion and zero-inflation.  Negative-binomial and Tweedie distributions are common choices to model the distribution of counts [[citation?]].
A key concern with the two-stage approach is the propogation of the uncertainty in the detectability estimates to the second-stage spatial model.  Early attempts to address this focused on bootstrapping [[citations]], but more recent work has pointed to potentially irresolvable difficulties  [[flesh out]] with the bootstrapping approach and instead proposed propagating error based on a second-order Taylor approximation of detectability around the maximum-likelihood estimate from the first stage \citep{bravington_reliable_2018-1}.

\citet{yuan_point_2017} presented a one-stage approach for analysing line-transect distance sampling within a Bayesian framework that simulateously estimates the detectability
and spatial distribution of animals.  Key to their approach is the use of sparse Gaussian Markov random field (GMRF) approximations to continously indexed Gaussian random fields (GRF) and a log-Gaussian Cox point process model that avoids the need to bin data into counts and naturally sidesteps issues of zero-inflation and dispersion.  The observation process is formulated as a thinning of the log-Gaussian Cox process.  Both the log-intensity of points and the detection function were modelled using a stochastic partial differential equation approach \citep{lindgren_explicit_2011}.  The model was specified as a latent Gaussian model and fitted using integrated nested Laplace approximations using the \texttt{R-INLA} package \citep{rue_approximate_2009}.

Here we present a similar approach to \citet{yuan_point_2017} but instead focus on the case of point transect data.  Point transects require special treatment to account for the fact that the area surveyed increases with increasing distance from the observer.  Our case study is, to the best of our knowledge, the first analysis of point transect distance sampling data as a thinned point process.  However, the point process viewpoint is not new and has been applied numerous times to line transect data [[citations here]].

@@

Standard distance sampling approaches specify $g(s)$ as a function that decays with increasing distance from the observer.  For example, if $r(s)$ denotes the distance of a point at $s$ from the observer, the half-normal thinning probability function is $g(s | \sigma) = \exp(-r(s)^2 / 2\sigma^2)$ where $\sigma$ is a variance parameter to be estimated using the observed distances.  The parameter $\sigma$ can only be estimated if some assumption is made about the intensity of the animal locations.  Without such an assumption detectability and intensity are confounded.  The standard assumption of distance sampling is that the intensity is constant with respect to changes in $r(s)$.  Thus any observed differences from uniformity can be attributed to detectability, and not variation in the intensity.

A point transect distance sampling survey consists of a set of $K$ sampled regions.  The $k$-th sampled region we denote $\Omega_k \subset \mathbb{R}^2$ and the total surveyed region $\Omega = \cup_{k=1}^K \Omega_k$.  For simplicity we assume that all sampled regions are discs with radius $W$ and $\Omega_j \cap \Omega_k = \emptyset$ for all $j \neq k$.  The thinning probability function $g(s)$ is then defined relative to the positions of the surveyed regions.  The probability of observing a point outside the surveyed region is zero.  We denote the probability of observing a point $s \in \Omega_k$ as $g_k(s)$.  Since the surveyed regions are non-overlapping each location $s$ is associated with a single thinning probability function $g_k$.  For example, for an observer at location $s_k \in \Omega_k$, the half-normal thinning probability function is $g_k(s) = \exp(-\lVert s - s_k \rVert_2^2 / 2\sigma^2)$. The assumption of non-overlapping survey regions can be relaxed by including extra information such as the time of each observation.  For simplicity we do not consider this here.  The thinning probability function for any $s \in \Omega$ is then given by $g(s) = g_{k(s)}(s)$ where $k(s) = k$ for $s \in \Omega_k$.

@@

The above specifies a model for all animal locations in the area of interest.  The log-intensity for observed points is $\log\tl(s) = \log\lambda(s) + \log g(s)$.  This presents a problem since $\log g(s)$ is not typically linear in its parameters. However, for any single set of parameter values, $g(s)$ is essentially an offset term.  That is to say, if we could specify $g(s)$ as known then we could proceed with inference with a standard additive linear predictor.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.7\linewidth]{figures/2002studyareapoints_pt}
	\caption{Study area (light blue) showing the 2002 survey points (black dots) in Hakalau Forest National Wildlife Refuge, \hawaii{} Island.}
	\label{fig:2002studyareapointspt}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.7\linewidth]{figures/2002counts_pt}
	\caption{Counts of \akepa{} by point during the 2002 survey.}
	\label{fig:2002countspt}
\end{figure}

@@

The CV plot shows a measure of uncertainty relative to the mean intensity in each prediction grid cell.  However, often we are interested in the uncertainty around the mean intensity on the scale of the intensity function itself - perhaps seeking to identify areas where we can confidently predict particularly high or low intensities.  To attempt to communicate this uncertainty around the mean it is common [[references?]] to present the lower and upper quantiles corresponding to a 95\% credible interval for each prediction cell (\autoref{fig:intensity-quantiles}).

@@

This difference becomes clear if one generates realizations of the posterior intensity field.  Three such realizations are in \autoref{fig:intensity-realizations}.  There are some notable differences when compared to the mean and quantile summary maps.  First, each realization has finer-grained spatial structure than the mean of the field.  What this means in practice is that we should expect more spatial structure in our collected data than what we might expect if we only look at the posterior mean of the field.  The mean field plot hides the spatial structure the model has been able to estimate from the data.  In general, the mean of a random field will contain less spatial structure than realisations of the field.

@@

The quantile plots are more concerning.  As mentioned above, the temptation is to interpret these as possible intensities that produced our observed data.  However, they in fact show the quantiles of the \textit{marginal} probability density for each prediction location.  This essentially treats each prediction cell as independent of the others. Whilst any one prediction location may be well described by such quantiles, to present them side by side in one image obscures the fact that the probability that all (or some subset of) the prediction cells would \textit{simultaneously} achieve that value will be in fact much lower than the quantile value and in many cases would be essentially zero.  To demonstrate this we treated the lower and upper quantile plots as though they were intensities and integrated them to obtain an expected abundance estimate of approximately 2500 for the 0.025 quantile and 12,300 for the 0.975.  A naive (and tempting) interpretation of these numbers is as upper and lower limits of the 95\% credible interval for abundance. However, these abundance estimates are essentially impossible according to the estimated posterior for abundance (see \autoref{fig:realized-abundance-posterior}).  This emphasises that the temptation to treat these maps as showing possible intensity surfaces leads to outcomes (and possible management decisions) that are inconsistent with the very model that generated the maps.

As an alternative we suggest using excursion sets and excursion functions.  This is a recently approach to representing uncertainty in stochastic processes that specifically considers
joint probability of events across multiple locations.  The positive excursion set with level $u$ for a function $f(s)$ with domain $\Omega$ is $A_u^{+}(f) = \{ s \in \Omega ; f(s) < u \}$.  For a stochastic process $\lambda(s)$ the positive excursion set with level $u$ and probability $1 - \alpha$ is

@@

Excursion sets can be estimated by considering increasing candidate sets $D$ and a sequential integration scheme to estimate probabilities.  An implementation is available in the \texttt{excursions} package available through the Comprehensive R Archive Network.

@@

  A key feature of these plots is that they display either discrete sub-regions of the area of interest (for excursion sets) or the probability of membership of such discrete sub-regions (for excursion functions).  We believe the risk of interpreting these maps as intensities is much lower than for the quantile plots.  The fact that it is possible to estimate discrete regions may also be useful to communicate results to conservation managers and decision-makers.  We anticipate the excursion function will be most useful in communicating to audiences with statistical training.

@@

TODO HERE:  Something about how this is all a one-stage model and all these plots are based on sampling from the joint posterior of all parameters.  As such all plots can be viewed as incorporating the uncertainty from the detection model.  This is a lovely and natural way to present results and an advantage of having a joint model.  I'm not sure varprop can do this?

@@

\begin{enumerate}
	\item Note the model-based Bayesian hierarchical approach provides a framework for working with distance sampling data.  This allows simultaneous estimation of parameters associated with the detection process and the spatial distribution.  It is in principle extendible... somehow...
	\item Note that Bayesian distance sampling has been done before but that the INLA approach avoids data augmentation and MCMC.
	\item Iterated INLA is a powerful and flexible way to include non-linear components to a predictor and has applications beyond this application to distance sampling
	\item More to do in future to test how many parameters can be included in the non-linear component and what types of non-linear components are feasible

>>>>

% RJC 09/08/2019 Added co-authors and affiliations. Note that I used \affil[]{} instead of \address[]{}
\author[1,*]{Andrew E Seaton}
\author[1,2]{Richard J Camp}
\author[3]{Finn Lindgren}
\author[1]{Janine B Illian}
\author[1]{David L Borchers}
\author[1]{David L Miller}      % Ask about co-authoring the manuscript
\author[1]{Len Thomas}          % Ask about co-authoring the manuscript
\author[1]{Stephen T Buckland}  % Ask about co-authoring the manuscript
\author[4]{Steve J Kendall}     % It is Steve's data we are using.

%\address{Centre for Research into Ecological \& Environmental Modelling and School of Mathematics \& Statistics, University of St Andrews, St Andrews, Fife, Scotland}
\affil[1]{Centre for Research into Ecological \& Environmental Modelling and School of Mathematics \& Statistics, University of St Andrews, St Andrews, Fife, Scotland}
\affil[2]{U. S. Geological Survey, Pacific Island Ecosystems Research Center, P.O. Box 44, \hawaii{} National Park, HI 96718, U.S.A.}
\affil[3]{School of Mathematics, University of Edinburgh, Edinburgh, Scotland}
\affil[4]{U. S. Fish and Wildlife, Big Island National Wildlife Refuge Complex, 60 Nowelo St., Suite 100, Hilo, HI  96720, U.S.A.}
\affil[*]{Correspondence: Andrew E Seaton, Email: aes22@st-andrews.ac.uk}

@@

Introduction outline:
\begin{enumerate}
	\item Estimation of abundance and spatial distribution of animals is an important topic
	\item Distance sampling is a family of methods that aim to do this whilst accounting for imperfect detectability of animals
	\item Distance sampling began life under a design-based inference paradigm
	\item Randomised survey design is used in two ways.  Firstly, to justify the assumption that the true distribution of animals is uniform with respect to distance from the observer.  Secondly, to construct Horvitz-Thompson-like estimators of animal density, which are then converted to abundance estimates for regions of interest
	\item Interest in the spatial distribution of animals has led to the development of approaches that combine traditional distance sampling methods and spatially explicit generalized additive models
	\item In this approach the estimated detection probabilities for discrete spatial units are used as an offset in a generalized additive model where the response variable is a count of detections within each unit
	\item This two-stage approach needs a way to propagate uncertainty from the detection model to the spatial model
	\item Recently [[JOYCE PAPER]] presented a one-stage approach that simultaneously estimates the detection and spatial processes, avoids binning data into counts and removes the need to consider uncertainty propagation
	\item This approach considered line-transect distance sampling as thinned log-Gaussian Cox process
	\item To estimate the detection function [[JOYCE et al]] constructed a spline-like function using an SPDE approach and fit the model using \texttt{R-INLA}
	\item In this paper we show how to achieve a similar one-stage model for point transect distance sampling
	\item Point transect surveys require a different approach than line transects because the area surveyed increases with distance from observer
	\item In addition to accounting for this [RJC - what does 'this' refer to?] we present an alternative way to specify the detection function.  We replace the spline approach with a parametric family of detection functions such as half-normal and hazard rate functions [RJC - I thought we couldn't use the hazard rate (hr) function? If the updated inlabru package includes the hr detection function model I need to rerun dsm as hr better fits the data.]
	\item  These detection functions are typically not linear in their parameters and so cannot be straightforwardly estimated using methods for additive linear predictors
	\item To solve this we implement an iterative approach based on a Taylor approximation of the detection function and fit the model using the \texttt{inlabru} package which extends the functionality of \texttt{R-INLA}
	\item This represents a novel development in distance sampling methods that allows simultaneous estimation of parametric detection functions and the spatial distribution of animals from point transect data
\end{enumerate}

The estimation of the size of wild populations of animals is a critical objective within ecology and conservation \citep{schwarz_estimating_1999}.  Distance sampling methods aim to achieve this by using a spatially explicit sampling design and estimating the detectability of animals as a function of distance from observer \citep{buckland_advanced_2008, buckland_distance_2015}.  Standard distance sampling approaches use a hybrid of design- and model-based inference to estimate population size.  Probability of detection is modelled as a function of distance from observer and a randomized sampling design allows the construction of Horvitz-Thompson-like estimators of animal density \citep{horvitz_generalization_1952,  buckland_advanced_2008}.

Over time interest grew in fully model-based approaches which allow the use of non-randomized survey designs by considering a spatial model for animal density so observations can be associated with spatially indexed covariates and abundance can, in principle, be estimated for any subregion within the study area \citep{johnson_model-based_2010, miller_spatial_2013, buckland_model-based_2016}.  Model-based distance sampling has typically been implemented in a two-stage modelling process.  In the first stage, detectability is estimated within each sampling unit.  In the second stage, detections within sampling units are binned into counts and used as the response variable in a generalized additive model framework where the detectability estimates from the first stage are used as an offset in the second stage.  Due to often sparse nature of wildlife survey data this may require consideration of over or under-dispersion and zero-inflated distributions.  Negative-binomial and Tweedie distributions are common choices to model the distribution of counts \citep{wood_gam_2017}.
A key concern with the two-stage approach is the propagation of the uncertainty in the detectability estimates to the second-stage spatial model.  Early attempts to address this focused on bootstrapping \citep{lahiri_resampling_2003}, but more recent work with the bootstrapping approach has pointed to potentially irresolvable difficulties that require resampling blocks at multiple spatial scales. Instead \cite{bravington_reliable_2018-1} proposed propagating error based on a second-order Taylor approximation of detectability around the maximum-likelihood estimate from the first stage.

\citet{yuan_point_2017} presented a one-stage approach for analysing line-transect distance sampling within a Bayesian framework that simultaneously estimates the detectability and spatial distribution of animals.  Key to their approach is the use of sparse Gaussian Markov random field (GMRF) approximations to continuously indexed Gaussian random fields (GRF) and a log-Gaussian Cox point process model that avoids the need to bin data into counts and naturally sidesteps issues of zero-inflation and dispersion.  The observation process is formulated as a thinning of the log-Gaussian Cox process.  Both the log-intensity of points and the detection function were modelled using a stochastic partial differential equation approach \citep{lindgren_explicit_2011}.  The model was specified as a latent Gaussian model and fitted using integrated nested Laplace approximations using the \texttt{R-INLA} package \citep{rue_approximate_2009}.

Here we present a similar approach to \citet{yuan_point_2017} but instead focus on the case of point-transect distance sampling data.  Point transects require special treatment to account for the fact that the area surveyed increases with increasing distance from the observer.  Our case study is, to the best of our knowledge, the first analysis of point-transect distance sampling data as a thinned point process.  However, the point process viewpoint is not new and has been applied numerous times to line transect data [[citations here]].

@@

Standard distance sampling approaches specify $g(s)$ as a function that decays with increasing distance from the observer.  For example, if $r(s)$ denotes the distance of a point at $s$ from the observer, the half-normal thinning probability function is $g(s | \sigma) = \exp(-r(s)^2 / 2\sigma^2)$ where $\sigma$ is a variance parameter to be estimated using the observed distances.  The parameter $\sigma$ can only be estimated if an assumption is made about the intensity of the animal locations.  Without such an assumption detectability and intensity are confounded.  The standard assumption of distance sampling is that the intensity is constant with respect to changes in $r(s)$.  Thus any observed differences from uniformity can be attributed to detectability, and not variation in the intensity.

A point transect distance sampling survey consists of a set of $K$ sampled regions.  The $k$-th sampled region we denote $\Omega_k \subset \mathbb{R}^2$ and the total surveyed region $\Omega = \cup_{k=1}^K \Omega_k$.  For simplicity we assume that all sampled regions are discs with radius $W$ and $\Omega_j \cap \Omega_k = \emptyset$ for all $j \neq k$.  The thinning probability function $g(s)$ is then defined relative to the positions of the surveyed regions.  We denote the probability of observing a point $s \in \Omega_k$ as $g_k(s)$.  The probability of observing a point outside the surveyed region is zero.  [RJC - I switched the order of the two previous sentences. That way we start with the relevant probability, followed by defining the probability elsewhere is zero]  Since the surveyed regions are non-overlapping each location $s$ is associated with a single thinning probability function $g_k$.  For example, for an observer at location $s_k \in \Omega_k$, the half-normal thinning probability function is $g_k(s) = \exp(-\lVert s - s_k \rVert_2^2 / 2\sigma^2)$. The assumption of non-overlapping survey regions can be relaxed by including extra information such as the time of each observation.  For simplicity we do not consider this here.  The thinning probability function for any $s \in \Omega$ is then given by $g(s) = g_{k(s)}(s)$ where $k(s) = k$ for $s \in \Omega_k$.

@@

The above specifies a model for all animal locations in the area of interest.  The log-intensity for observed points is $\log\tl(s) = \log\lambda(s) + \log g(s)$.  This presents a problem since $\log g(s)$ is not typically linear in its parameters. However, for any single set of parameter values, $g(s)$ is essentially an offset term.  That is, if we specify $g(s)$ as known then we can proceed with inference within a standard additive linear predictor framework.

@@

\begin{figure}
	\centering
	\includegraphics[scale=0.5]{figures/2002studyareapoints_pt}
	\caption{Study area (light blue) showing the 2002 survey points (black dots) in Hakalau Forest National Wildlife Refuge, \hawaii{} Island.}
	\label{fig:2002studyareapointspt}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[scale=0.5]{figures/2002counts_pt}
	\caption{Counts of \akepa{} by point during the 2002 survey.}
	\label{fig:2002countspt}
\end{figure}

@@

The CV plot shows a measure of uncertainty relative to the mean intensity in each prediction grid cell.  However, often we are interested in the uncertainty around the mean intensity on the scale of the intensity function itself - perhaps seeking to identify areas where we can confidently predict particularly high or low intensities.  To attempt to communicate this uncertainty around the mean it is common [[references?]] to present the lower and upper quantiles corresponding to a 95\% credible interval for each prediction cell (\autoref{fig:intensity-quantiles}). [RJC - I don't think a reference is needed to substantiate presenting credible intervals or computing them using quantiles, particularly in a stats journal.]

@@

This difference becomes clear if one generates realizations of the posterior intensity field.  Three such realizations are in \autoref{fig:intensity-realizations}.  There are some notable differences when compared to the mean and quantile summary maps.  First, each realization has finer-grained spatial structure than the mean of the field.  In practice, this indicates that we should expect more spatial structure in our collected data than what we might expect if we only look at the posterior mean of the field.  The mean field plot hides the spatial structure the model has been able to estimate from the data.  In general, the mean of a random field will contain less spatial structure than realisations of the field. [RJC - 'First' is not followed up with subsequent ordinal numbers; I suggest rewriting the above sentence to remove 'first' or add ordinal numbers below.]

@@

The quantile plots are more concerning.  As mentioned above, the temptation is to interpret these as possible intensities that produced our observed data.  However, they in fact show the quantiles of the \textit{marginal} probability density for each prediction location.  This essentially treats each prediction cell as independent of the others. Whilst any one prediction location may be well described by such quantiles, to present them side by side in one image obscures the fact that the probability that all (or some subset of) the prediction cells would \textit{simultaneously} achieve that value will be in fact much lower than the quantile value and in many cases would be essentially zero.  [RJC - I stumbled over this sentence several times and I think it needs to be rewritten.]  To demonstrate this we treated the lower and upper quantile plots as though they were intensities and integrated them to obtain an expected abundance estimate of approximately 2500 for the 0.025 quantile and 12,300 for the 0.975.  A naive (and tempting) interpretation of these numbers is as upper and lower limits of the 95\% credible interval for abundance. However, these abundance estimates are essentially impossible according to the estimated posterior for abundance (see \autoref{fig:realized-abundance-posterior}).  This emphasises that the temptation to treat these maps as showing possible intensity surfaces leads to outcomes (and possible management decisions) that are inconsistent with the very model that generated the maps.

As an alternative, we suggest using excursion sets and excursion functions.  This is a recently developed approach to representing uncertainty in stochastic processes that specifically considers joint probability of events across multiple locations.  The positive excursion set with level $u$ for a function $f(s)$ with domain $\Omega$ is $A_u^{+}(f) = \{ s \in \Omega ; f(s) < u \}$.  For a stochastic process $\lambda(s)$ the positive excursion set with level $u$ and probability $1 - \alpha$ is

@@

Excursion sets can be estimated by considering increasing candidate sets $D$ and a sequential integration scheme to estimate probabilities.  An implementation is available in the \texttt{excursions} package \citep{bolin_calculating_2018} available through the Comprehensive R Archive Network \citep{r_2017}.

@@

A key feature of these plots is that they display either discrete sub-regions of the area of interest (for excursion sets) or the probability of membership of such discrete sub-regions (for excursion functions).  We believe the risk of interpreting these maps as intensities is much lower than for the quantile plots.  The fact that it is possible to estimate discrete regions may also be useful to communicate results to conservation practitioners, land and resource managers and policy decision-makers.  We anticipate the excursion function will be most useful in communicating to audiences with statistical training. [RJC - I think statistically savvy audiences will readily grasp the excursion functions. Do you mean "without" statistical training?]

@@

TODO HERE:  Something about how this is all a one-stage model and all these plots are based on sampling from the joint posterior of all parameters.  As such all plots can be viewed as incorporating the uncertainty from the detection model.  This is a lovely and natural way to present results and an advantage of having a joint model.  I'm not sure varprop can do this? [RJC - Varprop does do this through the posterior covariance V_beta matrix of the intercept, smoother and detection probability terms. But it doesn't do so in such an elegant way. Varprop is still reliant on a two-stage model.]

@@

\begin{enumerate}
	\item Note the model-based Bayesian hierarchical approach provides a framework for working with distance sampling data.  This allows simultaneous estimation of parameters associated with the detection process and the spatial distribution.  It is in principle extendible... somehow...
	\item Note that Bayesian distance sampling [RJC - References for this include Oedekoven et al (2014, JABES 19:219-239) and Buckland et al. (2015). We could also include Buckland et al. (2016, JABES 21:58-75) but it deals more with model-based design distance sampling than Bayesian approaches specifically. The more recent papers that reference Cornelia's paper are applied. It is also worth looking at Yuan's paper on point process models of blue whales.] has been done before but that the INLA approach avoids data augmentation and MCMC.
	\item Iterated INLA is a powerful and flexible way to include non-linear components to a predictor and has applications beyond this application to distance sampling
	\item More to do in future to test how many parameters can be included in the non-linear component and what types of non-linear components are feasible
